---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FACT - (Feature Attributions for Clustering)

<!-- badges: start -->

<!-- badges: end -->

To get value from a clustering algorithm, it is important to understand the assignment procedure of an algorithm that assigns instances to clusters. FACT is an algorithm agnostic framework that provides feature attribution while preserving the integrity of the data.

## Features

- SMART (Scoring Metric After Permutation) permutes feature sets to measure the 
  sensitivity of algorithms to changes in cluster assignments.
- IDEA (Isolated Effect on Assignment) visualises local and global changes in
  cluster assignments over one- and two-dimensional feature spaces.

## Installation

You can install the development version of FACT like so:

``` r
# Development version
remotes::install_github("henrifnk/FACT")
```

## Quickstart

We want to divide American states by their standardized crime rates in 3
clusters.

```{r example}
library(FACT)
library(mlr3cluster)
attributes_scale = attributes(scale(USArrests))
```

```{r, echo=FALSE}
knitr::kable(head(scale(USArrests)), caption = "USArrests Data Set", digits = 2)
```

Therefore, we use a c-means algorithm from `mlr3cluster`.

```{r}
tsk_usa = TaskClust$new(id = "usarest", backend = data.frame(scale(USArrests)))
c_lrn = lrn("clust.cmeans", centers = 3, predict_type = "prob")
c_lrn$train(tsk_usa)
```

Then, we create a `ClustPredictor` that wraps all the information needed
for our methods.

```{r}
predictor = ClustPredictor$new(c_lrn, data = tsk_usa$data(), y = c_lrn$model$membership)
```

How does `Assault` effect the clustering outcome created by the
`c-means`?

```{r, results='hide', fig.keep='all', message = FALSE, out.width="65%", fig.align='center'}
idea_assault = IDEA$new(predictor, "Assault", grid.size = 50)
idea_assault$plot_globals(0.5)
```

# Citation

If you use FACT in a scientific publication, please
[cite](https://arxiv.org/abs/2209.10578) it as:

    Scholbeck, C. A., Funk, H., & Casalicchio, G. (2022). Algorithm-Agnostic Interpretations for Clustering. arXiv preprint arXiv:2209.10578.

BibTeX:

```{tex}
@article{FACT_22,
  title={Algorithm-Agnostic Interpretations for Clustering},
  author={Scholbeck, Christian A and Funk, Henri and Casalicchio, Giuseppe},
  journal={arXiv preprint arXiv:2209.10578},
  year={2022}
}
```
