---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FACT - (Feature Attributions for Clustering)

<!-- badges: start -->

<!-- badges: end -->

To get value from a clustering algorithm, it is important to understand the assignment procedure of an algorithm that assigns instances to clusters. `FACT` is an algorithm agnostic framework that provides feature attribution while preserving the integrity of the data.

## Features

- `SMART` (Scoring Metric After Permutation) permutes feature sets to measure the 
  sensitivity of algorithms to changes in cluster assignments.
- `IDEA` (Isolated Effect on Assignment) visualises local and global changes in
  cluster assignments over one- and two-dimensional feature spaces.

## Installation

You can install the development version of `FACT` like so:

``` r
# Development version
remotes::install_github("henrifnk/FACT")
```

## Quickstart

We want to divide American states by their standardized crime rates in 3
clusters.

```{r example}
library(FACT)
library(mlr3cluster)
attributes_scale = attributes(scale(USArrests))
```

```{r, echo=FALSE}
knitr::kable(head(scale(USArrests)), caption = "USArrests Data Set", digits = 2)
```

Therefore, we use a c-means algorithm from `mlr3cluster`.

```{r}
tsk_usa = TaskClust$new(id = "usarest", backend = data.frame(scale(USArrests)))
c_lrn = lrn("clust.cmeans", centers = 3, predict_type = "prob")
c_lrn$train(tsk_usa)
```

Then, we create a `ClustPredictor` that wraps all the information needed
for our methods.

```{r}
predictor = ClustPredictor$new(c_lrn, data = tsk_usa$data(), y = c_lrn$model$membership)
```

How does `Assault` effect the partitions created by `c-means` clustering?

The `sIDEA` plot shows:

- **x-Axis**: The domain in the feature space of `Assault` were realizations of
observations can be found (visualised by the `geom_rug`).
- **y-Axis**: The associated soft labels score of cluster k, `f(k)`.
- **solid line**: The estimated marginal, global effect of a cluster k throughout feature space.
- **transparent area**: 50\% of the mass of individual effects. This area plots the
variance in effects throughout feature space.

```{r, results='hide', fig.keep='all', message = FALSE, out.width="65%", fig.align='center'}
idea_assault = IDEA$new(predictor, "Assault", grid.size = 50)
idea_assault$plot_globals(0.5)
```

**Short Interpretation:**

- States in cluster 1 (red) are marginally associated with the lowest `Assault` rate.
- States in cluster 3 (blue) are marginally associated with a relatively low `Assault` rate.
- States in cluster 2 (green) are marginally associated with a relatively high `Assault` rate.

# Citation

If you use `FACT` in a scientific publication, please
[cite](https://arxiv.org/abs/2209.10578) it as:

    Scholbeck, C. A., Funk, H., & Casalicchio, G. (2022). Algorithm-Agnostic Interpretations for Clustering. arXiv preprint arXiv:2209.10578.

BibTeX:

```{tex}
@article{FACT_22,
  title={Algorithm-Agnostic Interpretations for Clustering},
  author={Scholbeck, Christian A and Funk, Henri and Casalicchio, Giuseppe},
  journal={arXiv preprint arXiv:2209.10578},
  year={2022}
}
```
